---
title: "create_bootstrap_lmer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create_bootstrap_samples}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```


```{r setup }
library(boots.lmer)
```


This vignette contains a walk-through of `boots.lmer::boots_samples` and `boots.lmer::boots_lmer` function.



# boots_samples 
The `boots_samples` function returns a list of bootstrap index and subject names, made especially for longitudinal data. Longitudinal data formatted where multiple observations should be chosen cautiously when sampling for boostrap:

1. sampling unit should be subject. That is, if one subject is chosen in the boostrap sample, the subject's entire observations should be included. 
2. the returning subject's name should be different if they were selected more than once, to use for random effects. 

Here's the example. The below example can be found in `help(boots_samples)` page as well.

<br>
<br>

# Generate example data

Here's an example data:
```r
set.seed(1249)
subjects<-c("Sarah","John","Beth","Anna","Chris","Blake")
example.subject<-sample(subjects, size = 50,replace = TRUE)
example.dat<-data.frame("Y"=rnorm(n=length(example.subject)),
                        "X1"=rpois(n=length(example.subject), lambda = 3),
                        "X2"=rnorm(n=length(example.subject)),
                        "X3"=rbeta(n=length(example.subject), shape1 = 3, shape2 = 0.5),
                        "subjects"=example.subject)
usethis::use_data(example.dat)
```

We have saved this data, so we will load it to use it. 

```{r}
data(example.dat, package="boots.lmer")
knitr::kable(head(example.dat), caption = "Table 1. Example data set.")
```

The example dataset contains total 50 observations with some outcome of interests (`Y`), 3 covariates (`X1, X2, X3`), and 6 subjects (Sarah, John, Beth, Anna, Chris, and Blck) that were randomly sampled. 




# Creating bootstrap samples using a longitudinal data


Here's `boots_samples` output, if we use `example.dat$subjects` as our subject ID to create 5000 bootstrap samples.:


```r
set.seed(128)
output<-boots.samples(dat = example.dat, sub.id = "subjects",B=5000)
usethis::use_data(output, overwrite = TRUE)
```
To make sure that the rest of explanation makes sense, we have save the data and loaded it. 

```{r}
data(output, package="boots.lmer")

knitr::kable(head(output[[1]]),caption = "Table 2. Four boostrap samples of example.dat data sets. The sampling unit is subject, hence the number of rows of each data set is expected to be different from one another.",row.names = FALSE)
```

The output is a list of 4, where each list conatins a `data.table` object with column `index` and `no.repeat.id`. The `index` column gives the row number of the oriianl data set (`example.dat`) to select as bootstrap sample. The `no.repeat.sub.id` column returns the bootstrap selected subjects from the `example.dat$subject` column, with trailing `__#`. The `#` shows the number of times the subject was included in the bootstrap sampling. 

For example, 
```{r}
unique((output[[1]])$no.repeat.id)
```

In the first bootstrap sample, Beth and Sarah have been sampled once, and Chris and Blake have been sampled twice. 



<br>
<br>

# Fit boots_lmer function


`boots_lmer` function requires at least 3 inputs. First, we need to input a string the outcome variable for `y`, and a vector of characters of covarites `X` as they appear in our data set `example.dat`. We then have to input the original data in `dat`, which is example.dat. We can also input number of cores (`num_workers`) and user-defined LMER formula (`formula`) if desires. The default number of cores to use is 2 and default formula is additive model of all covariates on outcome, with the random effect of the subject returned from the `boots_samples` function. 

```r
set.seed(4484)
lmer_out<-boots_lmer(y="Y", X=c("X1","X2","X3"), dat=example.dat, boots.samples.list = output)
usethis::use_data(lmer_out, overwrite = TRUE)
```

Here's the summary:

```{r}
data(lmer_out, package="boots.lmer")
summary(lmer_out)
```


Here's the plot:

```{r}
plot(lmer_out)
```


