---
title: "create_lmer_fits"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create_lmer_fits}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

```{r setup}
library(boots.lmer)
```

This vignette shows how to generate LMER fits using the `boots.lmer::boots.lmer` function from the bootstrap samples generated with the `boots.lmer::boots.samples` function. Please refer to 'create_bootstrap_samples' vignette to see detailed description of how to use `boots.samples` function.

# Generate example data and bootstrap samples

Here's an example data and its 4 bootstrap samples.:
```{r , eval=FALSE}
set.seed(1020)
example.subject<-c("Sarah","John","Beth","Anna","Sarah","Sarah","Chris","Blake","John","Anna")
example.dat<-data.frame("Y"=rnorm(n=length(example.subject)),
                        "X1"=rpois(n=length(example.subject), lambda = 3),
                        "X2"=rnorm(n=length(example.subject)),
                        "X3"=rbeta(n=length(example.subject), shape1 = 3, shape2 = 0.5),
                        "subjects"=example.subject) #generate example data


output<-boots.samples(dat=example.dat,sub.id ="subjects",B=4) #a list of 4 bootstrap samples 

```

